---
title: "README"
author: "Amanda Elert"
date: "4/13/2020"
output: md_document
---

# Loading the libraries needed
```{r}
rm(list = ls())

library(data.table)
library(tidyverse)
library(ggrepel)
library(broom)
library(randomForest)
```

# Background about the data
Our project covers [NFL play-by-play data from 2009 to 2018 from Kaggle](https://www.kaggle.com/maxhorowitz/nflplaybyplay2009to2016), and we are using betting data from Kaggle (link to data here) to contextualize the play-by-play data.
<br>
<br>
The [extra point yard line was moved](http://www.nfl.com/news/story/0ap3000000493347/article/nfl-moves-extra-point-to-15yard-line-for-2015-season
) from the 2-yardline to the 15-yardline in 2015 because the NFL wanted to increase the number of 2-point conversion attempts to boast ratings. We want to analyze how attempting more 2-point conversions, especially in the 4th quarter, can benefit teams.

# Data Cleaning
[Follow this link for using an open source Git extension to deal with versioning large files.](https://git-lfs.github.com/) This is how we were able to load the large play-by-play dataset.
```{r}
nfl_data <-  read_csv("NFLPlaybyPlay 2009-2018.csv")
spread_data = read.csv("spreadspoke_scores.csv")
spread_data$newdate <- strptime((spread_data$schedule_date), "%m/%d/%Y")
spread_data$newdate = format(spread_data$newdate, "%Y-%m-%d")
spread_data$newdate = as.Date(as.character(spread_data$newdate))

spread_data$home_team_abbv <- ifelse(spread_data$team_home=="Arizona Cardinals", "ARI",
                                     ifelse(spread_data$team_home=="Atlanta Falcons", "ATL", 
                                            ifelse(spread_data$team_home=="Baltimore Ravens", "BAL",
                                                   ifelse(spread_data$team_home=="Buffalo Bills", "BUF",
                                                          ifelse(spread_data$team_home=="Carolina Panthers", "CAR",
                              ifelse(spread_data$team_home=="Chicago Bears", "CHI",
                                      ifelse(spread_data$team_home=="Cincinnati Bengals", "CIN",
                                            ifelse(spread_data$team_home=="Cleveland Browns", "CLE",
                                                   ifelse(spread_data$team_home=="Dallas Cowboys", "DAL",
                                                          ifelse(spread_data$team_home=="Denver Broncos", "DEN", 
                              ifelse(spread_data$team_home=="Detroit Lions", "DET",
                                     ifelse(spread_data$team_home=="Green Bay Packers", "GB",
                                            ifelse(spread_data$team_home=="Houston Texans", "HOU",
                                                   ifelse(spread_data$team_home=="Indianapolis Colts", "IND",
                              ifelse(spread_data$team_home=="Jacksonville Jaguars", "JAX",
                                     ifelse(spread_data$team_home=="Kansas City Chiefs", "KC",
                                            ifelse(spread_data$team_home=="Los Angeles Rams", "LA",
                                                   ifelse(spread_data$team_home=="Los Angeles Chargers", "LAC",
                              ifelse(spread_data$team_home=="Miami Dolphins", "MIA",
                                     ifelse(spread_data$team_home=="Minnesota Vikings", "MIN",
                                            ifelse(spread_data$team_home=="New England Patriots", "NE",
                                                ifelse(spread_data$team_home=="New Orleans Saints", "NO",
                              ifelse(spread_data$team_home=="New York Giants", "NYG",
                                     ifelse(spread_data$team_home=="New York Jets", "NYJ",
                                            ifelse(spread_data$team_home=="Oakland Raiders", "OAK",
                                                   ifelse(spread_data$team_home=="Philadelphia Eagles", "PHI",
                              ifelse(spread_data$team_home=="Pittsburgh Steelers", "PIT",
                                     ifelse(spread_data$team_home=="San Diego Chargers", "SD",
                                            ifelse(spread_data$team_home=="Seattle Seahawks", "SEA",
                              ifelse(spread_data$team_home=="San Francisco 49ers", "SF",
                                     ifelse(spread_data$team_home=="St. Louis Rams", "STL",
                                            ifelse(spread_data$team_home=="Tampa Bay Buccaneers", "TB",
                              ifelse(spread_data$team_home=="Tennessee Titans", "TEN",
                                     ifelse(spread_data$team_home=="Washington Redskins", "WAS", "NA"
                                            ))))))))))))))))))))))))))))))))))

#Jacksonville's abbreviation changed from JAC to JAX in 2013
nfl_data$home_team <- ifelse(nfl_data$home_team == "JAC", "JAX", nfl_data$home_team)

colnames(spread_data)[19] <- "home_team"
colnames(spread_data)[18] <- "game_date"
glimpse(spread_data)

nfl_data <- nfl_data %>% 
                left_join(spread_data, by=c("game_date", "home_team"))

post_td_plays <- nfl_data %>% 
  select(play_id,
         game_id,
         home_team,
         away_team,
         posteam,
         posteam_type,
         defteam,
         posteam_score,
         defteam_score,
         score_differential,
         posteam_score_post,
         defteam_score_post,
         score_differential_post,
         desc,
         game_date,
         qtr,
         game_seconds_remaining,
         play_type,
         yards_gained,
         contains("two_point"),
         kicker_player_name,
         kicker_player_id,
         blocked_player_id,
         blocked_player_name,
         contains("extra_point"),
         wp,
         def_wp,
         home_wp,
         away_wp,
         wpa,
         home_wp_post,
         away_wp_post,
         ydsnet, 
         ep, 
         epa,
         shotgun,
         qb_dropback,
         team_favorite_id, 
         spread_favorite, 
         over_under_line, 
         stadium,
         stadium_neutral, 
         weather_temperature,
         weather_wind_mph,
         weather_humidity,
         weather_detail, 
         score_differential_post)%>% 
  mutate(year = substr(game_id, 1, 4)) %>%
  filter(extra_point_attempt == 1 | two_point_attempt == 1 | defensive_extra_point_attempt == 1 | defensive_two_point_attempt == 1)

post_td_plays$home_team <- as.factor(as.character(post_td_plays$home_team))
post_td_plays$away_team <- as.factor(as.character(post_td_plays$away_team))
post_td_plays$team_favorite_id <- as.factor(as.character(post_td_plays$team_favorite_id))

#Create one column that separates the types of extra point(s) try
post_td_plays <- post_td_plays %>% 
  mutate(extra_point_type = ifelse(extra_point_attempt == 1,
                               "Kick", "Two-PointConversion"))

#str(post_td_plays)
post_td_plays$posteam <- as.factor(post_td_plays$posteam)
post_td_plays$defteam <- as.factor(post_td_plays$defteam)
post_td_plays$play_type <- as.factor(post_td_plays$play_type)
post_td_plays$two_point_attempt <- as.factor(post_td_plays$two_point_attempt)
post_td_plays$two_point_conv_result <- as.factor(post_td_plays$two_point_conv_result)
post_td_plays$kicker_player_id <- as.factor(post_td_plays$kicker_player_id)
post_td_plays$extra_point_attempt <- as.factor(post_td_plays$extra_point_attempt)
post_td_plays$extra_point_result <- as.factor(post_td_plays$extra_point_result)
post_td_plays$year <- as.factor(post_td_plays$year)
post_td_plays$weather_humidity <- as.numeric(as.character(post_td_plays$weather_humidity))
post_td_plays$extra_point_type <- as.factor(as.character(post_td_plays$extra_point_type))
glimpse(post_td_plays)

#Dropping the no_play play type from the data...it's all penalties
post_td_plays <- post_td_plays %>% 
  filter(play_type != "no_play")
```

```{r}
# Ratio of XPT:2PT attempts
post_td_plays %>% 
  group_by(year, extra_point_attempt, two_point_attempt) %>% 
  summarise(n = n())

#Mean WPA change based on play type for post-touchdown 
post_td_plays %>% 
  group_by(play_type) %>% 
  summarise(mean(wpa, na.rm = T))
```

# Data Visualization
```{r}
# 2-pt Conversion Distribution
post_td_plays %>% 
  filter(two_point_attempt == 1) %>% 
  group_by(year, two_point_conv_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = two_point_conv_result)) +
  geom_line(aes(color = two_point_conv_result)) +
  geom_point(aes(color = two_point_conv_result)) +
  annotate(geom="text", x='2016', y=49, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of 2-pt Conversions in the NFL from 2009 to 2018",x="Year", y = "Count")

# XPT Conversion Distribution
post_td_plays %>% 
  filter(extra_point_attempt == 1) %>% 
  group_by(year, extra_point_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = extra_point_result)) +
  geom_line(aes(color = extra_point_result)) +
  geom_point(aes(color = extra_point_result)) +
  annotate(geom="text", x='2016', y=1100, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of Extra Points in the NFL from 2009 to 2018", x="Year", y = "Count")

#Graph 4th quarter win probability changes based on type of extra point try
post_td_plays %>% 
  filter(qtr == 4) %>% 
  ggplot(., aes(x=game_seconds_remaining, y=wpa, 
                                 color=extra_point_type)) + geom_point() + scale_x_reverse() +
  ggtitle("Win Probability Shifts in the 4th Quarter")

#Graph 4th quarter, 7 minutes left win probability changes based on type of extra point try
post_td_plays %>% 
  filter(qtr == 4, game_seconds_remaining <= 420) %>% 
  ggplot(., aes(x=game_seconds_remaining, y=wpa, 
                color=extra_point_type)) + geom_point() + scale_x_reverse()  +
  ggtitle("Win Probability Shifts in the last 7 min")
```

```{r}
#Success Rate of Extra Points (2015 and forward)
post_td_plays %>% 
  filter(extra_point_type == "Kick",
         year == "2015" | year == "2016" | year == "2017" | year == "2018") %>% 
  summarise(Extra_Point_Success = mean(extra_point_result == "good"))

#.9407 success rate

#Success Rate of Two Points
post_td_plays %>% 
  filter(extra_point_type == "Two-PointConversion") %>% 
  summarise(Two_Point_Success = mean(two_point_conv_result == "success"))
#.47345 success rate
```

# Data Modeling
```{r}
# mod1 <- lm(wpa ~ posteam + defteam + score_differential + play_type + two_point_attempt + two_point_conv_result + kicker_player_id + extra_point_attempt + extra_point_result + year, data = post_td_plays)

#Regression to see influential factors for two-point conversions specifcially
#Just a test, planning to further develop
lm_wpa <- post_td_plays %>%
  filter(wpa>0) %>% 
  lm(wpa ~ game_seconds_remaining + poly(score_differential, 2, raw=TRUE) + extra_point_type, data = .)

summary(lm_wpa)

```

```{r}
#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -15, extra_point_type = as.factor("Kick"))
kick_down15 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -15, extra_point_type = as.factor("Two-PointConversion"))
twopt_down15 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -14, extra_point_type = as.factor("Kick"))
kick_down14 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -14, extra_point_type = as.factor("Two-PointConversion"))
twopt_down14 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -13, extra_point_type = as.factor("Kick"))
kick_down13 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -13, extra_point_type = as.factor("Two-PointConversion"))
twopt_down13 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -12, extra_point_type = as.factor("Kick"))
kick_down12 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -12, extra_point_type = as.factor("Two-PointConversion"))
twopt_down12 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -11, extra_point_type = as.factor("Kick"))
kick_down11 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -11, extra_point_type = as.factor("Two-PointConversion"))
twopt_down11 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -10, extra_point_type = as.factor("Kick"))
kick_down10 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -10, extra_point_type = as.factor("Two-PointConversion"))
twopt_down10 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -9, extra_point_type = as.factor("Kick"))
kick_down9 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -9, extra_point_type = as.factor("Two-PointConversion"))
twopt_down9 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -8, extra_point_type = as.factor("Kick"))
kick_down8 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -8, extra_point_type = as.factor("Two-PointConversion"))
twopt_down8 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -7, extra_point_type = as.factor("Kick"))
kick_down7 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -7, extra_point_type = as.factor("Two-PointConversion"))
twopt_down7 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -6, extra_point_type = as.factor("Kick"))
kick_down6 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -6, extra_point_type = as.factor("Two-PointConversion"))
twopt_down6 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -5, extra_point_type = as.factor("Kick"))
kick_down5 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -5, extra_point_type = as.factor("Two-PointConversion"))
twopt_down5 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -4, extra_point_type = as.factor("Kick"))
kick_down4 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -4, extra_point_type = as.factor("Two-PointConversion"))
twopt_down4 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -3, extra_point_type = as.factor("Kick"))
kick_down3 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -3, extra_point_type = as.factor("Two-PointConversion"))
twopt_down3 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -2, extra_point_type = as.factor("Kick"))
kick_down2 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -2, extra_point_type = as.factor("Two-PointConversion"))
twopt_down2 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -1, extra_point_type = as.factor("Kick"))
kick_down1 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -1, extra_point_type = as.factor("Two-PointConversion"))
twopt_down1 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 0, extra_point_type = as.factor("Kick"))
kick_tied <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 0, extra_point_type = as.factor("Two-PointConversion"))
twopt_tied <- predict(lm_wpa, newdata = hypothetical_situation)

#Winning
#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 15, extra_point_type = as.factor("Kick"))
kick_up15 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 15, extra_point_type = as.factor("Two-PointConversion"))
twopt_up15 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 14, extra_point_type = as.factor("Kick"))
kick_up14 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 14, extra_point_type = as.factor("Two-PointConversion"))
twopt_up14 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 13, extra_point_type = as.factor("Kick"))
kick_up13 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 13, extra_point_type = as.factor("Two-PointConversion"))
twopt_up13 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 12, extra_point_type = as.factor("Kick"))
kick_up12 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 12, extra_point_type = as.factor("Two-PointConversion"))
twopt_up12 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 11, extra_point_type = as.factor("Kick"))
kick_up11 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 11, extra_point_type = as.factor("Two-PointConversion"))
twopt_up11 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 10, extra_point_type = as.factor("Kick"))
kick_up10 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 10, extra_point_type = as.factor("Two-PointConversion"))
twopt_up10 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 9, extra_point_type = as.factor("Kick"))
kick_up9 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 9, extra_point_type = as.factor("Two-PointConversion"))
twopt_up9 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 8, extra_point_type = as.factor("Kick"))
kick_up8 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 8, extra_point_type = as.factor("Two-PointConversion"))
twopt_up8 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 7, extra_point_type = as.factor("Kick"))
kick_up7 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 7, extra_point_type = as.factor("Two-PointConversion"))
twopt_up7 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 6, extra_point_type = as.factor("Kick"))
kick_up6 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 6, extra_point_type = as.factor("Two-PointConversion"))
twopt_up6 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 5, extra_point_type = as.factor("Kick"))
kick_up5 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 5, extra_point_type = as.factor("Two-PointConversion"))
twopt_up5 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 4, extra_point_type = as.factor("Kick"))
kick_up4 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 4, extra_point_type = as.factor("Two-PointConversion"))
twopt_up4 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 3, extra_point_type = as.factor("Kick"))
kick_up3 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 3, extra_point_type = as.factor("Two-PointConversion"))
twopt_up3 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 2, extra_point_type = as.factor("Kick"))
kick_up2 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 2, extra_point_type = as.factor("Two-PointConversion"))
twopt_up2 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 1, extra_point_type = as.factor("Kick"))
kick_up1 <- predict(lm_wpa, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 1, extra_point_type = as.factor("Two-PointConversion"))
twopt_up1 <- predict(lm_wpa, newdata = hypothetical_situation)

```

```{r}
#Merge Datasets
kick_wps <- rbind(kick_down15, kick_down14, kick_down13, kick_down12, kick_down11, kick_down10, kick_down9, kick_down8, kick_down7, kick_down6, kick_down5, kick_down4, kick_down3, kick_down2, kick_down1, kick_tied, kick_up1, kick_up2, kick_up3, kick_up4, kick_up5, kick_up6, kick_up7, kick_up8, kick_up9, kick_up10, kick_up11, kick_up12, kick_up13,kick_up14, kick_up15)

twopt_wps <- rbind(twopt_down15, twopt_down14, twopt_down13, twopt_down12, twopt_down11, twopt_down10,
                   twopt_down9, twopt_down8, twopt_down7, twopt_down6, twopt_down5, twopt_down4,
                   twopt_down3, twopt_down2, twopt_down1, twopt_tied, twopt_up1, twopt_up2, twopt_up3,
                  twopt_up4, twopt_up5, twopt_up6, twopt_up7, twopt_up8, twopt_up9, twopt_up10, 
                  twopt_up11, twopt_up12, twopt_up13, twopt_up14, twopt_up15)

wp_shifts <- cbind.data.frame(kick_wps, twopt_wps)    

row.names(wp_shifts) <- c("Down 15", "Down 14", "Down 13", "Down 12","Down 11", "Down 10",
                          "Down 9", "Down 8", "Down 7", "Down 6","Down 5", "Down 4",
                          "Down 3", "Down 2", "Down 1", "Tied","Up 1", "Up 2",
                          "Up 3", "Up 4", "Up 5", "Up 6", "Up 7", "Up 8", "Up 9", "Up 10",
                          "Up 11", "Up 12", "Up 13", "Up 14", "Up 15")

names(wp_shifts) <- c("Kick", "Two-Point")

print(wp_shifts)
```