---
title: "README"
author: "Amanda Elert"
date: "4/13/2020"
output: md_document
---

# Loading the libraries needed
```{r}
rm(list = ls())

library(data.table)
library(tidyverse)
library(ggrepel)
library(broom)
```

# Background about the data
Our project covers [NFL play-by-play data from 2009 to 2018 from Kaggle](https://www.kaggle.com/maxhorowitz/nflplaybyplay2009to2016), and we are using betting data from Kaggle (link to data here) to contextualize the play-by-play data.
<br>
<br>
The [extra point yard line was moved](http://www.nfl.com/news/story/0ap3000000493347/article/nfl-moves-extra-point-to-15yard-line-for-2015-season
) from the 2-yardline to the 15-yardline in 2015 because the NFL wanted to increase the number of 2-point conversion attempts to boast ratings. We want to analyze how attempting more 2-point conversions, especially in the 4th quarter, can benefit teams.

# Data Cleaning
[Follow this link for using an open source Git extension to deal with versioning large files.](https://git-lfs.github.com/) This is how we were able to load the large play-by-play dataset.
```{r}
nfl_data <-  read_csv("NFLPlaybyPlay 2009-2018.csv")
spread_data = read.csv("spreadspoke_scores.csv")
spread_data$newdate <- strptime((spread_data$schedule_date), "%m/%d/%Y")
spread_data$newdate = format(spread_data$newdate, "%Y-%m-%d")
spread_data$newdate = as.Date(as.character(spread_data$newdate))

spread_data$home_team_abbv <- ifelse(spread_data$team_home=="Arizona Cardinals", "ARI",
                                     ifelse(spread_data$team_home=="Atlanta Falcons", "ATL", 
                                            ifelse(spread_data$team_home=="Baltimore Ravens", "BAL",
                                                   ifelse(spread_data$team_home=="Buffalo Bills", "BUF",
                                                          ifelse(spread_data$team_home=="Carolina Panthers", "CAR",
                              ifelse(spread_data$team_home=="Chicago Bears", "CHI",
                                      ifelse(spread_data$team_home=="Cincinnati Bengals", "CIN",
                                            ifelse(spread_data$team_home=="Cleveland Browns", "CLE",
                                                   ifelse(spread_data$team_home=="Dallas Cowboys", "DAL",
                                                          ifelse(spread_data$team_home=="Denver Broncos", "DEN", 
                              ifelse(spread_data$team_home=="Detroit Lions", "DET",
                                     ifelse(spread_data$team_home=="Green Bay Packers", "GB",
                                            ifelse(spread_data$team_home=="Houston Texans", "HOU",
                                                   ifelse(spread_data$team_home=="Indianapolis Colts", "IND",
                              ifelse(spread_data$team_home=="Jacksonville Jaguars", "JAX",
                                     ifelse(spread_data$team_home=="Kansas City Chiefs", "KC",
                                            ifelse(spread_data$team_home=="Los Angeles Rams", "LA",
                                                   ifelse(spread_data$team_home=="Los Angeles Chargers", "LAC",
                              ifelse(spread_data$team_home=="Miami Dolphins", "MIA",
                                     ifelse(spread_data$team_home=="Minnesota Vikings", "MIN",
                                            ifelse(spread_data$team_home=="New England Patriots", "NE",
                                                ifelse(spread_data$team_home=="New Orleans Saints", "NO",
                              ifelse(spread_data$team_home=="New York Giants", "NYG",
                                     ifelse(spread_data$team_home=="New York Jets", "NYJ",
                                            ifelse(spread_data$team_home=="Oakland Raiders", "OAK",
                                                   ifelse(spread_data$team_home=="Philadelphia Eagles", "PHI",
                              ifelse(spread_data$team_home=="Pittsburgh Steelers", "PIT",
                                     ifelse(spread_data$team_home=="San Diego Chargers", "SD",
                                            ifelse(spread_data$team_home=="Seattle Seahawks", "SEA",
                              ifelse(spread_data$team_home=="San Francisco 49ers", "SF",
                                     ifelse(spread_data$team_home=="St. Louis Rams", "STL",
                                            ifelse(spread_data$team_home=="Tampa Bay Buccaneers", "TB",
                              ifelse(spread_data$team_home=="Tennessee Titans", "TEN",
                                     ifelse(spread_data$team_home=="Washington Redskins", "WAS", "NA"
                                            ))))))))))))))))))))))))))))))))))

#Jacksonville's abbreviation changed from JAC to JAX in 2013
nfl_data$home_team <- ifelse(nfl_data$home_team == "JAC", "JAX", nfl_data$home_team)

colnames(spread_data)[19] <- "home_team"
colnames(spread_data)[18] <- "game_date"
glimpse(spread_data)

nfl_data <- nfl_data %>% 
                left_join(spread_data, by=c("game_date", "home_team"))

post_td_plays <- nfl_data %>% 
  select(play_id,
         game_id,
         home_team,
         away_team,
         posteam,
         posteam_type,
         defteam,
         posteam_score,
         defteam_score,
         score_differential,
         posteam_score_post,
         defteam_score_post,
         score_differential_post,
         desc,
         game_date,
         qtr,
         game_seconds_remaining,
         play_type,
         yards_gained,
         contains("two_point"),
         kicker_player_name,
         kicker_player_id,
         blocked_player_id,
         blocked_player_name,
         contains("extra_point"),
         wp,
         def_wp,
         home_wp,
         away_wp,
         wpa,
         home_wp_post,
         away_wp_post,
         ydsnet, 
         ep, 
         epa,
         shotgun,
         qb_dropback,
         team_favorite_id, 
         spread_favorite, 
         over_under_line, 
         stadium,
         stadium_neutral, 
         weather_temperature,
         weather_wind_mph,
         weather_humidity,
         weather_detail)%>% 
  mutate(year = substr(game_id, 1, 4)) %>%
  filter(extra_point_attempt == 1 | two_point_attempt == 1 | defensive_extra_point_attempt == 1 | defensive_two_point_attempt == 1)

post_td_plays$home_team <- as.factor(as.character(post_td_plays$home_team))
post_td_plays$away_team <- as.factor(as.character(post_td_plays$away_team))
post_td_plays$team_favorite_id <- as.factor(as.character(post_td_plays$team_favorite_id))

#Create one column that separates the types of extra point(s) try
post_td_plays <- post_td_plays %>% 
  mutate(extra_point_type = ifelse(extra_point_attempt == 1,
                               "Kick", "Two-PointConversion"))

#str(post_td_plays)
post_td_plays$posteam <- as.factor(post_td_plays$posteam)
post_td_plays$defteam <- as.factor(post_td_plays$defteam)
post_td_plays$play_type <- as.factor(post_td_plays$play_type)
post_td_plays$two_point_attempt <- as.factor(post_td_plays$two_point_attempt)
post_td_plays$two_point_conv_result <- as.factor(post_td_plays$two_point_conv_result)
post_td_plays$kicker_player_id <- as.factor(post_td_plays$kicker_player_id)
post_td_plays$extra_point_attempt <- as.factor(post_td_plays$extra_point_attempt)
post_td_plays$extra_point_result <- as.factor(post_td_plays$extra_point_result)
post_td_plays$year <- as.factor(post_td_plays$year)
post_td_plays$weather_humidity <- as.numeric(as.character(post_td_plays$weather_humidity))
glimpse(post_td_plays)

#Dropping the no_play play type from the data...it's all penalties
post_td_plays <- post_td_plays %>% 
  filter(play_type != "no_play")
```

```{r}

# Ratio of XPT:2PT attempts
post_td_plays %>% 
  group_by(year, extra_point_attempt, two_point_attempt) %>% 
  summarise(n = n())

#Mean WPA change based on play type for post-touchdown 
post_td_plays %>% 
  group_by(play_type) %>% 
  summarise(mean(wpa, na.rm = T))
```

# Data Visualization
```{r}
# 2-pt Conversion Distribution
post_td_plays %>% 
  filter(two_point_attempt == 1) %>% 
  group_by(year, two_point_conv_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = two_point_conv_result)) +
  geom_line(aes(color = two_point_conv_result)) +
  geom_point(aes(color = two_point_conv_result)) +
  annotate(geom="text", x='2016', y=49, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of 2-pt Conversions in the NFL from 2009 to 2018",x="Year", y = "Count")

# XPT Conversion Distribution
post_td_plays %>% 
  filter(extra_point_attempt == 1) %>% 
  group_by(year, extra_point_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = extra_point_result)) +
  geom_line(aes(color = extra_point_result)) +
  geom_point(aes(color = extra_point_result)) +
  annotate(geom="text", x='2016', y=1100, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of Extra Points in the NFL from 2009 to 2018", x="Year", y = "Count")

#Graph 4th quarter win probability changes based on type of extra point try
post_td_plays %>% 
  filter(qtr == 4) %>% 
  ggplot(., aes(x=game_seconds_remaining, y=wpa, 
                                 color=extra_point_type)) + geom_point() + scale_x_reverse() +
  ggtitle("Win Probability Shifts in the 4th Quarter")

#Graph 4th quarter, 7 minutes left win probability changes based on type of extra point try
post_td_plays %>% 
  filter(qtr == 4, game_seconds_remaining <= 420) %>% 
  ggplot(., aes(x=game_seconds_remaining, y=wpa, 
                color=extra_point_type)) + geom_point() + scale_x_reverse()  +
  ggtitle("Win Probability Shifts in the last 7 min")
```

# Data Modeling
```{r}
# mod1 <- lm(wpa ~ posteam + defteam + score_differential + play_type + two_point_attempt + two_point_conv_result + kicker_player_id + extra_point_attempt + extra_point_result + year, data = post_td_plays)

#Regression to see influential factors for two-point conversions specifcially
#Just a test, planning to further develop
lm_twopoint <- post_td_plays %>% 
  lm(wp ~ game_seconds_remaining + score_differential + extra_point_type + weather_temperature + spread_favorite + over_under_line, data = .)

summary(lm_twopoint)
```

```{r}
#Success Rate of Extra Points (2015 and forward)
post_td_plays %>% 
  filter(extra_point_type == "Kick",
         year == "2015" | year == "2016" | year == "2017" | year == "2018") %>% 
  summarise(Extra_Point_Success = mean(extra_point_result == "good"))

#.9407 success rate

#Success Rate of Two Points
post_td_plays %>% 
  filter(extra_point_type == "Two-PointConversion") %>% 
  summarise(Two_Point_Success = mean(two_point_conv_result == "success"))
#.47345 success rate
```
```{r}
#10 minutes left, down 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -15, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down15 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 15, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -15, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down15 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 14, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -14, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down14 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 14, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -14, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down14 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 13, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -13, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down13 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 13, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -13, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down13 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -12, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down12 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 12, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -12, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down12 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 11, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -11, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down11 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 11, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -11, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down11 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 10, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -10, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down10 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 10, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -10, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down10 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 9, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -9, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down9 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 9, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -9, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down9 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 8, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -8, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down8 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 8, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -8, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down8 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 7, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -7, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down7 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 7, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -7, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down7 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 6, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -6, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down6 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 6, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -6, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down6 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 5, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -5, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down5 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 5, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -5, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down5 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 4, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -4, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down4 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 4, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -4, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down4 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 3, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -3, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down3 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 3, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -3, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down3 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 2, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -2, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down2 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 2, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -2, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down2 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 1, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -1, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_down1 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, down 1, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -1, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_down1 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, tied, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 0, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_tied <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, tied, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 0, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_tied <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 1, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 1, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up1 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 1, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 1, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up1 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 2, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 2, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up2 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 2, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 2, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up2 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 3, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 3, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up3 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 3, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 3, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up3 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 4, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 4, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up4 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 4, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 4, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up4 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 5, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 5, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up5 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 5, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 5, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up5 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 6, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 6, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up6 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 6, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 6, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up6 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 7, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 7, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up7 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 7, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 7, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up7 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 8, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 8, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up8 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 8, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 8, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up8 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 9, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 9, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up9 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 9, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 9, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up9 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 10, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 10, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up10 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 10, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 10, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up10 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 11, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 11, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up11 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 10, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 11, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up11 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 12, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 12, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up12 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 12, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 12, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up12 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 13, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 13, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up13 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 13, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 13, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up13 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 14, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 14, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up14 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 14, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 14, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up14 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 15, est change in probability for extra point 
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 15, extra_point_type = "Kick", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
kick_up15 <- predict(lm_twopoint, newdata = hypothetical_situation)

#10 minutes left, up 14, est change in probability for extra point
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = 15, extra_point_type = "Two-PointConversion", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)
twopt_up15 <- predict(lm_twopoint, newdata = hypothetical_situation)

kick_wps <- rbind(kick_down15, kick_down14, kick_down13, kick_down12, kick_down11, kick_down10, 
                  kick_down9, kick_down8, kick_down7,kick_down6, kick_down5, kick_down4, kick_down3,
                  kick_down2, kick_down1, kick_tied, kick_up1, kick_up2, kick_up3, kick_up4, kick_up5,
                  kick_up6, kick_up7, kick_up8, kick_up9, kick_up10, kick_up11, kick_up12, kick_up13,
                  kick_up14, kick_up15)

twopt_wps <- rbind(twopt_down15, twopt_down14, twopt_down13, twopt_down12, twopt_down11, twopt_down10, 
                   twopt_down9, twopt_down8, twopt_down7, twopt_down6, twopt_down5, twopt_down4,
                   twopt_down3, twopt_down2, twopt_down1, twopt_tied, twopt_up1, twopt_up2, twopt_up3,
                   twopt_up4, twopt_up5, twopt_up6, twopt_up7, twopt_up8, twopt_up9, twopt_up10, 
                   twopt_up11, twopt_up12, twopt_up13, twopt_up14, twopt_up15)

wp_shifts <- cbind.data.frame(kick_wps, twopt_wps)                
```

```{r}
# create new data frame with predictive data
hypothetical_situation <- data.frame(game_seconds_remaining = 600, score_differential = -4, play_type = "extra_point", weather_temperature = 60, spread_favorite = -6, over_under_line = 30)

# make predictions on wpa shift from going for 2pt
sims <- predict(lm_twopoint, newdata = hypothetical_situation)

mean(sims$.fitted)
#Expected to drop win probability by -0.00766
```

```{r}
#Finding some examples of game situations
post_td_plays %>% 
  filter(extra_point_type == "Two-PointConversion") %>% 
  arrange(wpa)

#2013-10-06 Cowboys vs Broncos, Cowboys 2pt Conversion 818 seconds remaining
cowboys_broncos_1006_2013 <- nfl_data %>% 
  filter(game_date == "2013-10-06", 
         home_team == "DAL")
#2011-11-20 Lions vs Panthers, Panthers 2pt Conversion 299 seconds remaining
lions_panthers_1120_2011 <- nfl_data %>% 
  filter(game_date == "2011-11-20", 
         home_team == "DET")

#2015-09-20 Bengals vs Chargers, Chargers 2pt Conversion Fails 260 seconds remaining
bengals_chargers_0920_2015 <- nfl_data %>% 
  filter(game_date == "2015-09-20", 
         home_team == "CIN")
#2017-10-29 Saints vs Bears, Bears 2pt Conversion Fails 238 seconds remaining
saints_bears_1029_2017 <- nfl_data %>% 
  filter(game_date == "2017-10-29", 
         home_team == "NO")

#Table showing between 210 and 270 seconds remaining, what coaches have done and how often successful
post_td_plays %>% 
  filter(game_seconds_remaining >=210 & game_seconds_remaining <= 270)

ggplot(saints_bears_1029_2017, aes(x=game_seconds_remaining, y=away_wp)) + geom_line() + xlim(0,300)

```

```{r}
library(shiny)

# User interface ----
ui <- fluidPage(
  titlePanel("NFL Two Point Conversions"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("YOu're the coach, choose what to do in several real NFL game situations"),
      
      selectInput("var", 
                  label = "Choose game",
                  choices = c("Percent White", "Percent Black",
                              "Percent Hispanic", "Percent Asian"),
                  selected = "Percent White"),
      
      sliderInput("range", 
                  label = "Range of interest:",
                  min = 0, max = 100, value = c(0, 100))
    ),
    
    mainPanel(plotOutput("map"))
  )
)

# Server logic ----
server <- function(input, output) {
  output$map <- renderPlot({
    data <- switch(input$var,
                   "Percent White" = counties$white,
                   "Percent Black" = counties$black,
                   "Percent Hispanic" = counties$hispanic,
                   "Percent Asian" = counties$asian)
    
    percent_map(var = data, color = "dodgerblue", legend.title = "Racial Makeup of the US by county", max = 100, min  = 0)
  })
}

# Run app ----
shinyApp(ui, server)
```
