---
title: "README"
author: "Amanda Elert"
date: "4/13/2020"
output: md_document
---

# Loading the libraries needed
```{r}
rm(list = ls())

library(data.table)
library(tidyverse)
library(ggrepel)
```

# Background about the data
Our project covers [NFL play-by-play data from 2009 to 2018 from Kaggle](https://www.kaggle.com/maxhorowitz/nflplaybyplay2009to2016), and we are using betting data from Kaggle (link to data here) to contextualize the play-by-play data.
<br>
<br>
The [extra point yard line was moved](http://www.nfl.com/news/story/0ap3000000493347/article/nfl-moves-extra-point-to-15yard-line-for-2015-season
) from the 2-yardline to the 15-yardline in 2015 because the NFL wanted to increase the number of 2-point conversion attempts to boast ratings. We want to analyze how attempting more 2-point conversions, especially in the 4th quarter, can benefit teams.

# Data Cleaning
[Follow this link for using an open source Git extension to deal with versioning large files.](https://git-lfs.github.com/) This is how we were able to load the large play-by-play dataset.
```{r}
nfl_data <- read_csv("NFLPlaybyPlay 2009-2018.csv")
spread_data = read.csv("spreadspoke_scores.csv")
spread_data$newdate <- strptime((spread_data$schedule_date), "%m/%d/%Y")
spread_data$newdate = format(spread_data$newdate, "%Y-%m-%d")
spread_data$newdate = as.character(spread_data$newdate)
post_td_plays <- nfl_data %>% 
  select(play_id,
         game_id,
         home_team,
         away_team,
         posteam,
         posteam_type,
         defteam,
         posteam_score,
         defteam_score,
         score_differential,
         posteam_score_post,
         defteam_score_post,
         score_differential_post,
         desc,
         game_date,
         play_type,
         yards_gained,
         contains("two_point"),
         kicker_player_name,
         kicker_player_id,
         blocked_player_id,
         blocked_player_name,
         contains("extra_point"),
         wp,
         def_wp,
         home_wp,
         away_wp,
         wpa,
         home_wp_post,
         away_wp_post) %>% 
  mutate(year = substr(game_id, 1, 4)) %>%
  filter(extra_point_attempt == 1 | two_point_attempt == 1 | defensive_extra_point_attempt == 1 | defensive_two_point_attempt == 1) %>% filter(home_team != "JAC", away_team != "JAC")

post_td_plays$newdate <- as.character(strptime(as.Date(post_td_plays$game_date), "%Y-%m-%d"))        
post_td_plays <- post_td_plays %>% left_join(spread_data, by = "newdate" )
post_td_plays$home_team <- as.character(post_td_plays$home_team)
post_td_plays$away_team <- as.character(post_td_plays$away_team)
post_td_plays$team_favorite_id <- as.character(post_td_plays$team_favorite_id)

post_td_plays <- post_td_plays %>% filter(home_team == team_favorite_id | away_team == team_favorite_id)


# Ratio of XPT:2PT attempts
post_td_plays %>% 
  group_by(year, extra_point_attempt, two_point_attempt) %>% 
  summarise(n = n())

```

# Data Visualization
```{r}
# 2-pt Conversion Distribution
post_td_plays %>% 
  filter(two_point_attempt == 1) %>% 
  group_by(year, two_point_conv_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = two_point_conv_result)) +
  geom_line(aes(color = two_point_conv_result)) +
  geom_point(aes(color = two_point_conv_result)) +
  annotate(geom="text", x='2016', y=49, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of 2-pt Conversions in the NFL from 2009 to 2018",x="Year", y = "Count")

# XPT Conversion Distribution
post_td_plays %>% 
  filter(extra_point_attempt == 1) %>% 
  group_by(year, extra_point_result) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n, group = extra_point_result)) +
  geom_line(aes(color = extra_point_result)) +
  geom_point(aes(color = extra_point_result)) +
  annotate(geom="text", x='2016', y=1100, label="2015: XPT moved from 2YL to 15YL", color="darkgray") +
  labs(title="Plot of Extra Points in the NFL from 2009 to 2018", x="Year", y = "Count")
```

# Data Modeling
```{r}
#str(post_td_plays)
post_td_plays$posteam <- as.factor(post_td_plays$posteam)
post_td_plays$defteam <- as.factor(post_td_plays$defteam)
post_td_plays$play_type <- as.factor(post_td_plays$play_type)
post_td_plays$two_point_attempt <- as.factor(post_td_plays$two_point_attempt)
post_td_plays$two_point_conv_result <- as.factor(post_td_plays$two_point_conv_result)
post_td_plays$kicker_player_id <- as.factor(post_td_plays$kicker_player_id)
post_td_plays$extra_point_attempt <- as.factor(post_td_plays$extra_point_attempt)
post_td_plays$extra_point_result <- as.factor(post_td_plays$extra_point_result)
post_td_plays$year <- as.factor(post_td_plays$year)

#str(post_td_plays)
#summary(post_td_plays)

# mod1 <- lm(wpa ~ posteam + defteam + score_differential + play_type + two_point_attempt + two_point_conv_result + kicker_player_id + extra_point_attempt + extra_point_result + year, data = post_td_plays)



```